Follow-up Samples

### 2017-01-03 submission/manuscript_outline_20161024.tex
update to get correct taxonomies

Colorectal cancer (CRC) continues to be a worldwide health problem with
early detection being used as a key component in mitigating deaths due
to the disease. Previous research suggests a link between stool
bacterial microbiome and CRC. In this study, we used a model based on
the microbiome, demographics, and prior medical history to classify
individuals as having a lesion (i.e.~adenoma or carcinoma). We then used
this model to characterize the change in the gut microbiota before and
after surgery. The overall objective was to investigate the changes in
the microbiome after surgery in patients with adenomas or carcinomas.
This model was tested on a 66 person group that included samples before
and after treatment to allow for the assessment of how the model adjusts
risk after treatment. The model used for prediction had an AUC of 0.763.
For the follow up samples our Random Forest model significantly
decreased the positive probability of lesion compared to the initial
samples for both adenoma (P-value = 3.64e-11) and carcinoma (P-value =
7.95e-08). Our model predicted that 36.4\% of the 67-person cohort had
normal colons and 63.6\% had a lesion. Some OTUs that changed the most
before and after treatment included OTUs that were affiliated with
members of Blautia, Clostridium\_XlVa and Escherichia/Shigella. Our
model suggests that treatment does significantly reduce the probability
of having a colonic lesion. Further surveillance of these individuals
will enable us to determine whether models such as the one we present
here can also be used to predict recurrence of colorectal cancer.

@@

\textbf{\emph{Most Important Variables to the Model}} Overall, there
were a total of 37 variables identified as being present in more than
50\% of the training models {[}Table S2{]}. The top 5 most important
bacterial OTUs were Bacteria (Otu000013), Escherichia/Shigella
(Otu000018), Bacteria (Otu000020), Ruminococcus (Otu000017), and
Porphyromonas (Otu000153). These 5 OTUs were present in at least 90 out
of the total 100 different 80/20 runs.

@@

\textbf{\emph{Specific OTUs in the Lesion Model are not Detected in
Follow Up Versus Initial Samples}} Overall, there were a total of 8 OTUs
that were common between the main lesion model and the model for
classifying initial and follow up samples specifically {[}Table S4{]}. A
total of 1 OTU was still significant after multiple comparison
correction and it's lowest taxonomic identification was to Blautia. In
general, Otu000012 (Blautia) was decreased from initial to follow up
{[}Figure 5{]}. The relative abundance was not drastically different
then the mean of the values observed in the control training set
{[}Figure 5{]}.

@@

\textbf{\emph{Analysis Overview}} Differences in FIT between initial and
follow ups for either adenoma or carcinoma were investigate. Next,
initial and follow up samples were analyzed for differences in alpha and
beta diversity. All OTUs used in the lesion model were also analyzed
using a paired wilcoxson test. The lesion model was then tested for
accuracy in prediction and whether it reduced the positive probability
of lesion after surgery. The most important OTUs for this were used to
build an updated model and this reduced feature model was assessed for
it's similarity to the original model. We then used the initial follow
up model to assess whether this model could classify samples better then
the lesion model. The most important OTUs were then identified from this
model and used to create a reduced feature initial follow up model. This
reduced feature model, as was done with the lesion model, was compared
to the full model for loss of accuracy. Finally, in order to investigate
the relative abundance of specific bacteria, that have been previously
associated with CRC, we selected OTUs that taxonomically classified to
Fusobacterium Nucleatum, Parvimonas Micra, Peptostreptococcus
Assacharolytica, and Porphyromonas Stomatis. Specifically, we wanted to
test if there were any differences based on whether the individual had
an adenoma or carcinoma.

>>>>

Colorectal cancer (CRC) continues to be a worldwide health problem with
early detection being used as a key component in mitigating deaths due
to the disease. Previous research suggests a link between stool
bacterial microbiome and CRC. In this study, we used a model based on
the microbiome, demographics, and prior medical history to classify
individuals as having a lesion (i.e.~adenoma or carcinoma). We then used
this model to characterize the change in the gut microbiota before and
after surgery. The overall objective was to investigate the changes in
the microbiome after surgery in patients with adenomas or carcinomas.
This model was tested on a 66 person group that included samples before
and after treatment to allow for the assessment of how the model adjusts
risk after treatment. The model used for prediction had an AUC of 0.763.
For the follow up samples our Random Forest model significantly
decreased the positive probability of lesion compared to the initial
samples for both adenoma (P-value = 3.64e-11) and carcinoma (P-value =
7.95e-08). Our model predicted that 36.4\% of the 67-person cohort had
normal colons and 63.6\% had a lesion. Some OTUs that changed the most
before and after treatment included OTUs that were affiliated with
members of Blautia, Streptococcus and Escherichia/Shigella. Our model
suggests that treatment does significantly reduce the probability of
having a colonic lesion. Further surveillance of these individuals will
enable us to determine whether models such as the one we present here
can also be used to predict recurrence of colorectal cancer.

@@

\textbf{\emph{Most Important Variables to the Model}} Overall, there
were a total of 37 variables identified as being present in more than
50\% of the training models {[}Table S2{]}. The top 5 most important
bacterial OTUs were Lachnospiraceae (Otu000013), Escherichia/Shigella
(Otu000018), Ruminococcaceae (Otu000020), Ruminococcus (Otu000017), and
Porphyromonas (Otu000153). These 5 OTUs were present in at least 90 out
of the total 100 different 80/20 runs.

@@

\textbf{\emph{Specific OTUs in the Lesion Model are not Detected in
Follow Up Versus Initial Samples}} Overall, there were a total of 11
OTUs that were common between the main lesion model and the model for
classifying initial and follow up samples specifically {[}Table S4{]}. A
total of 1 OTU was still significant after multiple comparison
correction and it's lowest taxonomic identification was to Blautia. In
general, Otu000012 (Blautia) was decreased from initial to follow up
{[}Figure 5{]}. The relative abundance was not drastically different
then the mean of the values observed in the control training set
{[}Figure 5{]}.

@@

\textbf{\emph{Analysis Overview}} Differences in FIT between initial and
follow ups for either adenoma or carcinoma were investigated. Next,
initial and follow up samples were analyzed for differences in alpha and
beta diversity. All OTUs used in the lesion model were also analyzed
using a paired wilcoxson test. The lesion model was then tested for
accuracy in prediction and whether it reduced the positive probability
of lesion after surgery. The most important OTUs for this were used to
build an updated model and this reduced feature model was assessed for
it's similarity to the original model. We then used the initial follow
up model to assess whether this model could classify samples better then
the lesion model. The most important OTUs were then identified from this
model and used to create a reduced feature initial follow up model. This
reduced feature model, as was done with the lesion model, was compared
to the full model for loss of accuracy. Finally, in order to investigate
the relative abundance of specific bacteria, that have been previously
associated with CRC, we selected OTUs that taxonomically classified to
Fusobacterium Nucleatum, Parvimonas Micra, Peptostreptococcus
Assacharolytica, and Porphyromonas Stomatis. Specifically, we wanted to
test if there were any differences based on whether the individual had
an adenoma or carcinoma.
